---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pii

<!-- badges: start -->
<!-- badges: end -->

The goal of pii is to flag columns that potentially contain personally identifiable information. This package was inspired by concerns that survey data might be shared without users realizing the files contain PII. It is based on a set of [standard guidlines](https://www.usaid.gov/sites/default/files/2022-05/508saa.pdf) from the United States Agency for International Development, though people can debate what is and isn't PII. 

## Installation

You can install the development version of pii from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jacobpstein/pii")
```

## Example

Let's say you're working with the `cars` data set. But this isn't just any version of `cars`, this is more like a data about characters from the hit feature film [*Cars*](https://cars.disney.com)! You happen to have phone numbers and the location of each cars house. A colleague at a partner org asks for the data, but you aren't sure if it has PII, so you run the `check_PII` function. 
```{r example}

library(dplyr)
library(pii)

# Set a seed for reproducibility
set.seed(101624)

# Number of rows in the cars dataset
n <- nrow(cars)

# Generate car phone numbers as strings
phone_numbers <- sprintf("555-%03d-%04d", sample(100:999, n, replace = TRUE), sample(1000:9999, n, replace = TRUE))

# Generate latitudes for where the cars live (range roughly between -90 and 90)
latitudes <- runif(n, min = -90, max = 90)

# Generate longitudes for where the cars live (range roughly between -180 and 180)
longitudes <- runif(n, min = -180, max = 180)

# Merge new columns into the cars dataset using mutate
cars_with_pii <- cars %>%
  mutate(PhoneNumber = phone_numbers,
         Latitude = latitudes,
         Longitude = longitudes)

# run our function over the data
cars_pii <- check_PII(cars_with_pii)

# take a look at the list
print(cars_pii)

```

The `check_PII` function flags combinations of columns that together could identify individuals. It also flags columns that contain names that suggest a column contains PII, like, `PhoneNumber.` 
